---
title: "CaseStudy2"
author: "Malcolm Carlson"
date: "12/3/2019"
output: html_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Executive Summary

### This case study explores an employee data set with 870 observations and 44 varialbles.  The first question of interest is what factors lead to employee attrition.  A prediction for employee income will also be provided based on a model.  Job specific trends will also be examined along with any other trends or data that may prove useful.


```{r}
library(dplyr)
library(magrittr)
library(tidyverse)
library(GGally)
library(corrplot)
library(DataExplorer)
library(class)
library(caret)
library(e1071)
library(investr)
library(MASS)
library(leaps)
library(Hmisc)
library(randomForest)
library(car)
```

# Import data and create numeric factors for non-numeric new categorical variables 
```{r}
attritdf<-read.csv(file="/Users/malco/Documents/MDS/MDS_6306_DDS/DDS-CaseStudy2/CaseStudy2-data.csv",header=TRUE)
noattritdf <- read.csv(file="/Users/malco/Documents/MDS/MDS_6306_DDS/DDS-CaseStudy2/CaseStudy2CompSetNoAttrition.csv",header=TRUE)
nosalarydf <- read.csv(file="/Users/malco/Documents/MDS/MDS_6306_DDS/DDS-CaseStudy2/CaseStudy2CompSetNoSalary.csv",header=TRUE)

attritdf$AgeGrp<-cut(attritdf$Age,breaks = c(17,24,34,44,54,Inf),labels=c('18-24','25-34','35-44','46-54','55+'))
noattritdf$AgeGrp<-cut(noattritdf$Age,breaks = c(17,24,34,44,54,Inf),labels=c('18-24','25-34','35-44','46-54','55+'))

attritdf <- attritdf %>% mutate(AttrNum = case_when(
  Attrition == "Yes" ~ 1,
  Attrition == "No" ~ 0
  ))

attritdf <- attritdf %>% mutate(GenNum = case_when(
  Gender == "Male" ~ 1,
  Gender == "Female" ~ 0
))
noattritdf <- noattritdf %>% mutate(GenNum = case_when(
  Gender == "Male" ~ 1,
  Gender == "Female" ~ 0
))

attritdf <- attritdf %>% mutate(TravLevel = case_when(
  BusinessTravel == "Non-Travel" ~ 0,
  BusinessTravel == "Travel_Rarely" ~ 1,
  BusinessTravel == "Travel_Frequently" ~ 2
))
noattritdf <- noattritdf %>% mutate(TravLevel = case_when(
  BusinessTravel == "Non-Travel" ~ 0,
  BusinessTravel == "Travel_Rarely" ~ 1,
  BusinessTravel == "Travel_Frequently" ~ 2
))


attritdf <- attritdf %>% mutate(OTNum = case_when(
  OverTime == "Yes" ~ 1,
  OverTime == "No" ~0
))
noattritdf <- noattritdf %>% mutate(OTNum = case_when(
  OverTime == "Yes" ~ 1,
  OverTime == "No" ~0
))

attritdf <- attritdf %>% mutate(MariStatNum = case_when(
  MaritalStatus == "Single" ~ 0,
  MaritalStatus == "Married" ~ 1,
  MaritalStatus == "Divorced" ~ 2
))
noattritdf <- noattritdf %>% mutate(MariStatNum = case_when(
  MaritalStatus == "Single" ~ 0,
  MaritalStatus == "Married" ~ 1,
  MaritalStatus == "Divorced" ~ 2
))

attritdf <- attritdf %>% mutate(EducNum = case_when(
  EducationField == "Medical" ~ 0,
  EducationField == "Life Sciences" ~ 1,
  EducationField == "Marketing" ~ 2,
  EducationField == "Technical Degree" ~ 3,
  EducationField == "Human Resources" ~ 4,
  EducationField == "Other" ~ 5
))
noattritdf <- noattritdf %>% mutate(EducNum = case_when(
  EducationField == "Medical" ~ 0,
  EducationField == "Life Sciences" ~ 1,
  EducationField == "Marketing" ~ 2,
  EducationField == "Technical Degree" ~ 3,
  EducationField == "Human Resources" ~ 4,
  EducationField == "Other" ~ 5
))

attritdf <- attritdf %>% mutate(JRoleNum = case_when(
  JobRole == "Research Director" ~ 0,
  JobRole == "Manufacturing Director" ~ 1,
  JobRole == "Sales Executive" ~ 2,
  JobRole == "Research Scientist" ~ 3,
  JobRole == "Sales Representative" ~ 4,
  JobRole == "Healthcare Representative" ~ 5,
  JobRole == "Manager" ~ 6,
  JobRole == "Laboratory Technician" ~ 7,
  JobRole == "Human Resources" ~ 8
))
noattritdf <- noattritdf %>% mutate(JRoleNum = case_when(
  JobRole == "Research Director" ~ 0,
  JobRole == "Manufacturing Director" ~ 1,
  JobRole == "Sales Executive" ~ 2,
  JobRole == "Research Scientist" ~ 3,
  JobRole == "Sales Representative" ~ 4,
  JobRole == "Healthcare Representative" ~ 5,
  JobRole == "Manager" ~ 6,
  JobRole == "Laboratory Technician" ~ 7,
  JobRole == "Human Resources" ~ 8
))       

allnumeric<-attritdf%>%dplyr::select(Age,AgeGrp,AttrNum,TravLevel,DailyRate,DistanceFromHome,Education,EducNum,EnvironmentSatisfaction,GenNum,HourlyRate,JobInvolvement,JobLevel,JRoleNum,JobSatisfaction,MariStatNum,MonthlyIncome,MonthlyRate,NumCompaniesWorked,OTNum,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)

```

# Convert integers to factors where values are categorical
```{r}

attritdf$JobInvolvement <- as.factor(attritdf$JobInvolvement)
noattritdf$JobInvolvement <- as.factor(noattritdf$JobInvolvement)
nosalarydf$JobInvolvement <- as.factor(nosalarydf$JobInvolvement)

attritdf$JobLevel <- as.factor(attritdf$JobLevel)
noattritdf$JobLevel <- as.factor(noattritdf$JobLevel)
nosalarydf$JobLevel <- as.factor(nosalarydf$JobLevel)

attritdf$JobSatisfaction <- as.factor(attritdf$JobSatisfaction)
noattritdf$JobSatisfaction <- as.factor(noattritdf$JobSatisfaction)
nosalarydf$JobSatisfaction <- as.factor(nosalarydf$JobSatisfaction)

attritdf$PerformanceRating <- as.factor(attritdf$PerformanceRating)
noattritdf$PerformanceRating <- as.factor(noattritdf$PerformanceRating)
nosalarydf$PerformanceRating <- as.factor(nosalarydf$PerformanceRating)

attritdf$RelationshipSatisfaction <- as.factor(attritdf$RelationshipSatisfaction)
noattritdf$RelationshipSatisfaction <- as.factor(noattritdf$RelationshipSatisfaction)
nosalarydf$RelationshipSatisfaction <- as.factor(nosalarydf$RelationshipSatisfaction)

attritdf$StockOptionLevel <- as.factor(attritdf$StockOptionLevel)
nosalarydf$StockOptionLevel <- as.factor(nosalarydf$StockOptionLevel)
noattritdf$StockOptionLevel <- as.factor(noattritdf$StockOptionLevel)

attritdf$TrainingTimesLastYear <- as.factor(attritdf$TrainingTimesLastYear)
noattritdf$TrainingTimesLastYear <- as.factor(noattritdf$TrainingTimesLastYear)
nosalarydf$TrainingTimesLastYear <- as.factor(nosalarydf$TrainingTimesLastYear)

attritdf$WorkLifeBalance <- as.factor(attritdf$WorkLifeBalance)
noattritdf$WorkLifeBalance <- as.factor(noattritdf$WorkLifeBalance)
nosalarydf$WorkLifeBalance <- as.factor(nosalarydf$WorkLifeBalance)

attritdf$Education <- as.factor(attritdf$Education)
noattritdf$Education <- as.factor(noattritdf$Education)
nosalarydf$Education <- as.factor(nosalarydf$Education)

attritdf$EnvironmentSatisfaction <- as.factor(attritdf$EnvironmentSatisfaction)
noattritdf$EnvironmentSatisfaction <- as.factor(noattritdf$EnvironmentSatisfaction)
nosalarydf$EnvironmentSatisfaction <- as.factor(nosalarydf$EnvironmentSatisfaction)

```

## EDA Section
### Analysis with DataExplorer library
```{r}
str(attritdf)
names(attritdf)
describe(attritdf)

plot_intro(attritdf)
plot_missing(attritdf)
plot_bar(attritdf)
plot_histogram(attritdf)
plot_qq(attritdf)

# Correlationapalooza
plot_correlation(allnumeric, type = c("all", "discrete", "continuous"),
  maxcat = 20L, cor_args = list(), geom_text_args = list(),
  title = NULL, ggtheme = theme_gray(),
  theme_config = list(legend.position = "bottom", axis.text.x =
  element_text(angle = 90)))

# Correlation with just continuous variables
attritcor1 <- attritdf%>%dplyr::select(NumCompaniesWorked,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,YearsAtCompany)
plot_correlation(attritcor1)

attrcont <- attritdf%>%dplyr::select(Attrition,Age,NumCompaniesWorked,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,YearsAtCompany,MonthlyIncome,TotalWorkingYears)
attrcont %>%
  filter(Attrition == "Yes") %>%
  select_if(is.numeric) %>%
  cor() %>%
  corrplot::corrplot()

attritdf %>% 
  dplyr::select(Age, DailyRate, DistanceFromHome, HourlyRate, MonthlyIncome, MonthlyRate, YearsAtCompany, YearsWithCurrManager, YearsSinceLastPromotion) %>% 
  gather(metric, value) %>% 
  ggplot(aes(value, fill = metric)) + 
  geom_density(show.legend = FALSE) + 
  facet_wrap(~ metric, scales = "free")

attritdf %>% 
  dplyr::select(Age, DailyRate, DistanceFromHome, HourlyRate, MonthlyIncome, MonthlyRate, YearsAtCompany, YearsWithCurrManager, YearsSinceLastPromotion) %>% 
  gather(metric, value) %>% 
  ggplot(aes(value, fill = metric)) + 
  geom_histogram(show.legend = FALSE) + 
  facet_wrap(~ metric, scales = "free")

```

## Scatterplot analysis
```{r}
plot_scatterplot(attritdf, by="MonthlyIncome")
plot_scatterplot(attritdf, by="YearsAtCompany")
plot_scatterplot(attritdf, by="YearsInCurrentRole")
plot_scatterplot(attritdf, by="YearsSinceLastPromotion")
plot_scatterplot(attritdf, by="YearsWithCurrManager")
plot_scatterplot(attritdf, by="NumCompaniesWorked")

```

# Explore AgeGrp
```{r}
attritdf%>%ggplot(aes(x=MonthlyIncome,YearsAtCompany,color=AgeGrp))+geom_point()
attritdf%>%ggplot(aes(x=MonthlyIncome,YearsAtCompany,color=AgeGrp))+geom_smooth()

attritdf%>%ggplot(aes(x=YearsSinceLastPromotion,PercentSalaryHike,color=AgeGrp))+geom_point()
attritdf%>%ggplot(aes(x=YearsSinceLastPromotion,PercentSalaryHike,color=AgeGrp))+geom_smooth()

attritdf%>%ggplot(aes(x=YearsWithCurrManager,MonthlyIncome,color=AgeGrp))+geom_point()
attritdf%>%ggplot(aes(x=YearsWithCurrManager,MonthlyIncome,color=AgeGrp))+geom_smooth()

attritdf%>%ggplot(aes(x=YearsAtCompany,MonthlyIncome,color=AgeGrp))+geom_point()
attritdf%>%ggplot(aes(x=YearsAtCompany,MonthlyIncome,color=AgeGrp))+geom_smooth()


# Attrition by Age Group
AgeGrpdf <- attritdf%>%dplyr::select(AgeGrp,Attrition)
AgeGrpdf <- AgeGrpdf%>%filter(Attrition=="Yes")
AgeGrpplot <- ggplot(AgeGrpdf,aes(x=AgeGrp,group=Attrition,color=AgeGrp))+geom_bar(aes(y = ..prop..,fill=factor(..x..)), stat="count")+ggtitle("Attrition by Age Group")+
  geom_text(aes(label = scales::percent(..prop..),y= ..prop.. ), stat= "count", vjust = -.1)+labs(y = "Percent")+ylab(NULL)+xlab("Age groups")

AgeGrpplot


```

## Attrition by JobRole
```{r}
AgeJLev <- attritdf%>%dplyr::select(JobLevel,Attrition)
AgeJLev <- AgeJLev%>%filter(Attrition=="Yes")

AgeJLevplot <- ggplot(AgeJLev,aes(x=JobLevel,group=Attrition,color=JobLevel))+geom_bar(aes(y = ..prop..,fill=factor(..x..)), stat="count")+ggtitle("Attrition by Job Level")+
geom_text(aes(label = scales::percent(..prop..),y= ..prop.. ), stat= "count", vjust = -.1)+labs(y = "Percent")+ylab(NULL)+xlab("Job Levels")

AgeJLevplot

```

## This section creates a dataframe with an abbreviated jobrole title jrole, and calculates the amount of attrition and non-attrition by job role.
```{r}

tempdf <- attritdf%>%dplyr::select(JobRole,Attrition,MonthlyIncome)%>%group_by(JobRole)
jrno <- tempdf%>%filter(Attrition=="Yes")
jrno <- jrno%>%summarise(AttrCnt=n())
jry <- tempdf%>%filter(Attrition=="No")
jry <- tempdf%>%summarise(NoAttrCnt=n())
AttrJobRoledf <- merge(jry,jrno)
AttrJobRoledf <- AttrJobRoledf %>% mutate(JRole = case_when(
  JobRole == "Healthcare Representative" ~ "HC_Rep",
  JobRole == "Human Resources" ~ "HR",
  JobRole == "Laboratory Technician" ~ "Lab_Tech",
  JobRole == "Manager" ~ "Mgr",
  JobRole == "Manufacturing Director" ~ "Manfact_Dir",
  JobRole == "Research Director" ~ "Re_Dir",
  JobRole == "Research Scientist" ~ "Re_Scientist",
  JobRole == "Sales Executive" ~ "SalesExec",
  JobRole == "Sales Representative" ~ "SalesRep"
))

```

## MonthlyIncome by JobRole
```{r}

ggplot(tempdf,aes(x=NULL,y=MonthlyIncome,fill=JobRole))+geom_boxplot()+ggtitle("Summary statistics by Monthly Income and Job Role")+ylab("Monthly Income")+xlab("Job Role")


attritdf%>%ggplot()+geom_bar(aes(y=JobSatisfaction,x=JobRole,fill=JobSatisfaction),stat="identity",size=4)+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ylab(NULL)+xlab(NULL)+ggtitle("Job Satisfaction Scores by Job Role")


```


## This section orders the data and reduces to top Job Roles with the most attrition.
```{r}

ggplot(AttrJobRoledf,aes(x=JobRole,y=AttrCnt))+geom_point()+ylab("Attrition Count")+xlab("Job Role")+theme(axis.text.x = element_text(angle = 90, hjust = 1))

# MonthlyIncome by job role
IncomeJobRolePlot <- ggplot(tempdf,aes(JobRole, MonthlyIncome,fill=JobRole))+geom_bar(stat = "identity",position = position_stack(reverse = TRUE))+coord_flip()
IncomeJobRolePlot


```

```{r}
# Continuous variable relationship analysis
# MonthlyIncome,YearsAtCompany,PercentSalaryHike,NumCompaniesWorked,TotalWorkingYears,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager

# Attrition by gender
AttritReduced <- attritdf%>%dplyr::select(Attrition,Age,AgeGrp,Gender,YearsSinceLastPromotion,YearsWithCurrManager,JobRole,)%>%group_by(Attrition="Yes")

AttritReduced%>%ggplot(aes(x=Gender,fill=Gender))+geom_bar()+ggtitle("Bar chart of Attrition by Gender")+xlab(NULL)+ylab(NULL)+geom_text(stat = "count",aes(label=..count..))+theme(legend.position = "none")

attritdf%>%ggplot(aes(x=Gender,fill=Gender))+geom_bar()+ggtitle("Bar chart by Gender")+xlab(NULL)+ylab(NULL)+geom_text(stat = "count",aes(label=..count..))+theme(legend.position = "none")

```


## Top factors that contribute to turnover multi linear regression 
## Test assumptions with residual plots and histogram of residuals. Inital qq plots were not as linear as I would have liked but appeared worse after log transforming.  I also looked at sqr the continuous variables but did not like how the data turned out.
```{r}

set.seed(123)
training.samples <- attritdf$AttrNum %>%
createDataPartition(p=0.8, list = FALSE)
train.data <- attritdf[training.samples, ]
test.data <- attritdf[-training.samples, ]

attritdflog <- attritdf%>%dplyr::select(MonthlyIncome,Attrition,AttrNum,Age,GenNum,YearsInCurrentRole,TotalWorkingYears,TrainingTimesLastYear,YearsSinceLastPromotion,JobSatisfaction,PerformanceRating,RelationshipSatisfaction,EnvironmentSatisfaction,DistanceFromHome,StockOptionLevel,NumCompaniesWorked,JobInvolvement,JRoleNum,EducNum,OTNum,TravLevel,MariStatNum)%>%mutate(logYSLP = log(YearsSinceLastPromotion),logDFH=log(DistanceFromHome),logNCW=log(NumCompaniesWorked),logTWY=log(TotalWorkingYears),sqrYSLP=(YearsSinceLastPromotion)^2)

# Initial model with most variables that made sense
model1<-lm(AttrNum~Age+GenNum+YearsInCurrentRole+TotalWorkingYears+TrainingTimesLastYear+YearsSinceLastPromotion+JobSatisfaction+PerformanceRating+RelationshipSatisfaction+EnvironmentSatisfaction+DistanceFromHome+StockOptionLevel+NumCompaniesWorked+JobInvolvement+JRoleNum+EducNum+OTNum+TravLevel+MariStatNum,data=attritdflog)

outlierTest(model1)
qqPlot(model1,main = "QQ Plot")
leveragePlots(model1,main = "Leverage Plots")
spreadLevelPlot(model1)
ncvTest(model1)
crPlots(model1)

mod1resid = rstudent(model1)
hist(mod1resid,col="red",main = "Histogram of residuals for initial model",ylab = NULL,xlab = NULL)

plot(model1)
summary(model1)
car::vif(model1)

# model after looking at p-values of multi-linear regression and their VIF scores
ggplot(attritdf,aes(sample=YearsSinceLastPromotion,colour=Attrition))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - YearsSinceLastPromotionw with Attrition")+xlab(NULL)+ylab(NULL)
#ggplot(attritdflog,aes(sample=logYSLP,colour=Attrition))+stat_qq()+stat_qq_line()

ggplot(attritdf,aes(sample=DistanceFromHome,colour=Attrition))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - DistanceFromHome with Attrition")+xlab(NULL)+ylab(NULL)
#ggplot(attritdflog,aes(sample=logDFH,colour=Attrition))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - log transformed DistanceFromHome with Attrition")

ggplot(attritdf,aes(sample=NumCompaniesWorked,colour=Attrition))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - NumCompaniesWorked with Attrition")+xlab(NULL)+ylab(NULL)
#ggplot(attritdflog,aes(sample=logNCW,colour=AttrNum))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - log transformed NumCompaniesWorked with Attrition")

ggplot(attritdflog,aes(sample=TotalWorkingYears,colour=Attrition))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - TotalWorkingYears with Attrition")+xlab(NULL)+ylab(NULL)
#ggplot(attritdflog,aes(sample=logTWY,colour=Attrition))+stat_qq()+stat_qq_line()+ggtitle("QQPlot - log transformed TotalWorkingYears with Attrition")


model2 <- lm(AttrNum~YearsSinceLastPromotion+JobSatisfaction+EnvironmentSatisfaction+DistanceFromHome+NumCompaniesWorked+JobInvolvement+OTNum, data=attritdflog)

summary(model2)
car::vif(model2)
plot(model2)
mod2resid = rstudent(model2)
hist(mod2resid,col="blue",main = "Histogram of residuals of final model",ylab = NULL,xlab = NULL)


```


## Other interesting trends
### Gender - Jobrole MonthlyIncome, Jobrole Attrition Jobrole NoAttrition
```{r}
attritdf%>%ggplot(aes(x=YearsAtCompany,MonthlyIncome,color=Gender))+geom_point()
attritdf%>%ggplot(aes(x=YearsAtCompany,MonthlyIncome,color=Gender))+geom_smooth()+ggtitle("Monthly income over years with company - M vs F")

attritdf%>%ggplot(aes(x=YearsWithCurrManager,MonthlyIncome,color=Gender))+geom_point()
attritdf%>%ggplot(aes(x=YearsWithCurrManager,MonthlyIncome,color=Gender))+geom_smooth()+ggtitle("Comparison of Income by Gender over time with same manager")
attritdf%>%ggplot(aes(x=Age,MonthlyIncome,color=Gender))+geom_point()
attritdf%>%ggplot(aes(x=Age,MonthlyIncome,color=Gender))+geom_smooth()+ggtitle("Monthly Income and Age - M vs F")

ggplot(attritdf,aes(sample=MonthlyIncome,colour=factor(Gender)))+stat_qq()+stat_qq_line()

genNCWviolin <- ggplot(attritdf,aes(x=JobRole,y=NumCompaniesWorked,fill=Gender))+geom_violin()+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ggtitle("Number of Companies worked by Job Role and Gender")
genNCWviolin
genYACviolin <- ggplot(attritdf, aes(x=JobRole, y=YearsAtCompany,fill=Gender))+geom_violin()+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ggtitle("Years At the Company by Job Role and Gender")
genYACviolin
genNCW <- ggplot(attritdf, aes(x=JobRole,y=MonthlyIncome,fill=Gender))+geom_col(position = "dodge2")+theme(axis.text.x = element_text(angle = 90, hjust = 1))
genNCW+ggtitle("Income by Job Role and Gender")


```


## Classifications
## knn (classify attrition as yes) noattritdf=testdata
```{r}
set.seed(123)
perc=.8
numks = 5

iterations=100
masterAcc = matrix(nrow = iterations, ncol = numks)
masterSensitivity = matrix(nrow = iterations)
masterSpecificity = matrix(nrow = iterations)

for(j in 1:iterations)
{
 accs = data.frame(accuracy=numeric(numks),k=numeric(numks))
 trainIndices = sample(1:dim(attritdf)[1],
 round(perc * dim(attritdf)[1]))
 train = attritdf[trainIndices,]
 test = attritdf[-trainIndices,]

  for(i in 1:numks)
  {
  classifications = knn(train[,c(2,12,15,16,18,22,25)],test[,c(2,12,15,16,18,22,25)],train$AttrNum,prob = TRUE, k = i)
  table(classifications,test$AttrNum)
  CM = confusionMatrix(table(classifications,test$AttrNum))
  masterAcc[j,i] = CM$overall[1]
  masterSensitivity[j]=CM$byClass[1]
  masterSpecificity[j]=CM$byClass[2]
  }
}
ptitle=print(paste0(numks," values of K for ",iterations," iterations."))
MeanAcc = colMeans(masterAcc)
MeanSpec = mean(masterSpecificity)
MeanSens = mean(masterSensitivity)
plot(seq(1,numks,1),MeanAcc, type = "l",xlab = "Values of K",ylab = "Accuracy",main = ptitle)
which.max(MeanAcc)

CM


```

## Naive Bayse Classification 
```{r}
set.seed(123)
perc=.8
numks = 5
vars = c("")
iterations=20
masterAcc = matrix(nrow = iterations, ncol = numks)
masterSensitivity = matrix(nrow = iterations)
masterSpecificity = matrix(nrow = iterations)
for(i in 1:iterations)
  {
  rn=sample(1:30,1)
  trainIndices = sample(1:dim(attritdf)[1],round(perc * dim(attritdf)[1]))
    train = attritdf[trainIndices,]
    test = attritdf[-trainIndices,]
  model = naiveBayes(train[,c(2,6,12,15,16,17,18,20,24,25,27,29,30,31,33,34)],as.factor(train$Attrition),laplace = 1)
  table(predict(model,test[,c(2,6,12,15,16,17,18,20,24,25,27,29,30,31,33,34)]),as.factor(test$Attrition))
  
  CM = confusionMatrix(table(predict(model,test[,c(2,6,12,15,16,17,18,20,24,25,27,29,30,31,33,34)]),as.factor(test$Attrition)))
  masterAcc[j]=CM$overall[1]
  masterSensitivity[i]=CM$byClass[1]
  masterSpecificity[i]=CM$byClass[1]
}
# Adjusted for noattritiondf
# c(2,5,11,14,15,16,17,19,23,24,26,28,29,30,32,33)

tmpdf <- attritdf[c(2,6,12,15,16,17,18,20,24,25,27,29,30,31,33,34)]
tmpdf2 <- noattritdf[c(2,5,11,14,15,16,17,19,23,24,26,28,29,30,32,33)]
attritOut <- predict(model,tmpdf2)
myout=cbind.data.frame(noattritdf$ID,attritOut)
colnames(myout) <- c("ID","Attrition")

write.csv(myout, file = "/Users/malco/Documents/MDS/MDS_6306_DDS/DDS-CaseStudy2/Case2PredictionsMalcolmCarlsonAttrition.csv",row.names = FALSE)
plot(attritOut,  col="darkgreen", main = "Plot of Attrition Predictions")
CM
```

## Regression analysis of salary
```{r}
set.seed(123)
training.samples <- attritdf$MonthlyIncome %>%
createDataPartition(p=0.8, list = FALSE)
train.data <- attritdf[training.samples, ]
test.data <- attritdf[-training.samples, ]

modelSalary<-lm(MonthlyIncome~JobLevel+JobRole+
TotalWorkingYears,data=attritdf)

pairs(~MonthlyIncome+JobLevel+TotalWorkingYears,data=attritdf)

plot(modelSalary)
car::vif(modelSalary)

mod1resid = rstudent(modelSalary)
hist(mod1resid,col="darkgreen",main = "Histogram of residuals for model",ylab = NULL,xlab = NULL)

predictions <- modelSalary %>% predict(test.data)
data.frame( R2 = R2(predictions, test.data$MonthlyIncome),
           RMSE = RMSE(predictions, test.data$MonthlyIncome),
            MAE = MAE(predictions, test.data$MonthlyIncome))

summary(modelSalary)
salaryPred <- modelSalary%>% predict(nosalarydf)
myout <- cbind.data.frame(nosalarydf$ID,salaryPred)
colnames(myout) <- c("ID","MonthlyIncome")
write.csv(myout, file = "/Users/malco/Documents/MDS/MDS_6306_DDS/DDS-CaseStudy2/Case2PredictionsMalcolmCarlsonSalary.csv",row.names = FALSE)


```



## Stepwise model
```{r}
attritdf2 <- attritdf%>%dplyr::select(Age,DailyRate,Department,DistanceFromHome,Education,EducationField,EnvironmentSatisfaction,HourlyRate,JobInvolvement,JobLevel,JobRole,JobSatisfaction,MaritalStatus,MonthlyIncome,MonthlyRate,NumCompaniesWorked,OverTime,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StandardHours,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)

# Fit the full model
set.seed(123)
full.model <- lm(MonthlyIncome ~., data = attritdf2)
# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)

pairs(~MonthlyIncome+JobLevel+TotalWorkingYears,data=attritdf)

plot(step.model)
car::vif(step.model)

step.modelid = rstudent(step.model)
hist(step.modelid,col="darkgreen",main = "Histogram of residuals for model",ylab = NULL,xlab = NULL)

predictions <- step.model %>% predict(test.data)
data.frame( R2 = R2(predictions, test.data$MonthlyIncome),
           RMSE = RMSE(predictions, test.data$MonthlyIncome),
            MAE = MAE(predictions, test.data$MonthlyIncome))

summary(step.model)
car::vif(step.model)

salaryPred <- step.model%>% predict(nosalarydf)
myout <- cbind.data.frame(nosalarydf$ID,salaryPred)
colnames(myout) <- c("ID","MonthlyIncome")
write.csv(myout, file = "/Users/malco/Documents/MDS/MDS_6306_DDS/DDS-CaseStudy2/Case2PredictionsMalcolmCarlsonSalary.csv",row.names = FALSE)

```

